What's in a name? On the automated topic naming of software maintenance activities

### 2010-03-06 icpc/icpc2010.tex
Desperate attempt to make it all work

@@

\section{Background}

@@

\subsection{Definitions}
We will use the following terminology in the rest of the paper.
A \emph{message} is a block of text written by a developer. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when a developer commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e., a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top ten most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.

@@

\subsection{Topic and Concept Extraction}
Topic extraction, sometimes called concept extraction, uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
independent word distributions (topics) from
	documents (commit log comments). Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining software repositories, in particular analyzing source code, bugs or developer comments.
\begin{comment}
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
\end{comment}

@@

\subsection{Supervised learning}
While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results. In supervised learning, the data is divided into slices. One slice is manually annotated by the domain expert, and the classifications he/she determines are applied to the remaining slices. In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test supervised learning.

\section{Methodology}

@@

\subsection{Generating the data}
From each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work. An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. We extracted these messages and indexed them by creation time. We summarized each message as a word distribution but removed stop-words such as common English words \emph{the} and \emph{at}.

@@

\subsection{Associating labels}
\begin{comment}
% We would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period.
\end{comment}

@@

\subsection{Supervised learning}

@@

\section{Observations and evaluation}

\subsection{Word list similarity}

@@

In general, this approach did not work out well as common labels dominated the less common labels. The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific. Table \ref{tbl:wordlist} lists results. A \emph{named   topic} is a topic with a matching label. There are \{periods X 20\} topics per project as we told LDA to extract 20 topics per period. All experiments were run on MaxDB 7.500 and MySQL 3.23 data.

%\XXX{this table should have mysql}
\begin{table}
	\centering
\begin{tabular}{c|c|c|c}
\toprule
Measure &		 \textsf{exp1} &	 \textsf{exp2} &	 \textsf{exp3} \\
\midrule
Topics &			500		 &			500  	 & 		500  \\
Named topics &		281      &			125      &		328  \\
Unnamed topics &	139      &			295      &      92   \\
\bottomrule
\end{tabular}
	\caption{Automatic topic labelling for MaxDB 7.500}
	\label{tbl:wordlist}

\end{table}

%This is the counts

@@

\subsection{Analysis of the unsupervised labelling}

@@

%\subsection{Classifying topics}
\subsection{Analysis of the supervised labelling}
\label{sec:suplabelling}

@@

\subsection{Applying multiple labels to topics}

@@

\subsection{Summary of techniques}

@@

\subsection{Visualization}

@@

\begin{figure*}
  \centering
  \includegraphics[height=.9\textheight]{small-maxdb-time-smear-plot-Equal}
  \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
  \label{fig:maxdb}
\end{figure*}

\begin{figure*}[h]
  \centering
  \includegraphics[width=.99\textwidth]{small-mysql-time-smear-plot-Equal}
  \caption{MySQL 3.23: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
  \label{fig:mysql}
\end{figure*}

\subsection{Comparing MaxDB and MySQL}

@@

\paragraph{Do label frequencies change over time?} Yes, MySQL's label
frequencies decreased as it got older. Usability and reliability
labels became more and more infrequent as it matured. Maintainability topics
became more prevalent as MaxDB matured.

\paragraph{Do the different projects differ in their relative topic
  interest?} Yes they do, MySQL 3.23 had more proportionally more
functionality labelled topics while MaxDB had proportionally more
efficiency and portability related topics.

@@

\subsection{Annotation observations}

@@

\subsection{Effectiveness}

@@

\subsection{Threats to validity}

@@

\section{Related work}

@@

\section{Conclusions and future work}

@@

\section{Acknowledgements and appendix}
Our data and scripts are available at \url{http://softwareprocess.es/nomen/}

>>>>

\newcommand{\shrink}{}

@@

\shrink
\section{Background}

@@

\shrink
\subsection{Definitions}
We will use the following terminology in the rest of the paper.
A \emph{message} is a block of text written by a developer. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when a developer commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e., a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top ten most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.

@@

\shrink
\subsection{Topic and Concept Extraction}
Topic extraction, sometimes called concept extraction, uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
independent word distributions (topics) from
	documents (commit log comments). Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining software repositories, in particular analyzing source code, bugs or developer comments.
\begin{comment}
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
\end{comment}

@@

\shrink
\subsection{Supervised learning}
While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results. In supervised learning, the data is divided into slices. One slice is manually annotated by the domain expert, and the classifications he/she determines are applied to the remaining slices. In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test supervised learning.

\shrink 
\section{Methodology}

@@

\shrink
\subsection{Generating the data}
From each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work. An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. We extracted these messages and indexed them by creation time. We summarized each message as a word distribution but removed stop-words such as common English words \emph{the} and \emph{at}.

@@

\shrink
\subsection{Associating labels}
\begin{comment}
% We would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period.
\end{comment}

@@

\shrink
\subsection{Supervised learning}

@@

\shrink
\section{Observations and evaluation}

\shrink
\subsection{Word list similarity}

@@

In general, this approach did not work out well as common labels
dominated the less common labels. The related words for
\emph{correctness}, for example, tended to be too lengthy and
non-specific. 
%Table \ref{tbl:wordlist} lists results. 
A \emph{named   topic} is a topic with a matching label. There are \{periods X 20\} topics per project as we told LDA to extract 20 topics per period. All experiments were run on MaxDB 7.500 and MySQL 3.23 data.

\begin{comment}
%\XXX{this table should have mysql}
\begin{table}
	\centering
\begin{tabular}{c|c|c|c}
\toprule
Measure &		 \textsf{exp1} &	 \textsf{exp2} &	 \textsf{exp3} \\
\midrule
Topics &			500		 &			500  	 & 		500  \\
Named topics &		281      &			125      &		328  \\
Unnamed topics &	139      &			295      &      92   \\
\bottomrule
\end{tabular}
	\caption{Automatic topic labelling for MaxDB 7.500}
	\label{tbl:wordlist}

\end{table}
\end{comment}
%This is the counts

@@

\shrink
\subsection{Analysis of the unsupervised labelling}

@@

%\subsection{Classifying topics}
\shrink
\subsection{Analysis of the supervised labelling}
\label{sec:suplabelling}

@@

\shrink
\subsection{Applying multiple labels to topics}

@@

\shrink
\subsection{Summary of techniques}

@@

\shrink
\subsection{Visualization}

@@

\begin{figure*}
  \centering
  \includegraphics[height=.55\textheight]{small-maxdb-time-smear-plot-Equal}
  \caption{MaxDB 7.500: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
  \label{fig:maxdb}
\end{figure*}

\begin{figure*}%[h]
  \centering
  \includegraphics[width=.97\textwidth]{small-mysql-time-smear-plot-Equal}
  \caption{MySQL 3.23: Labelled topics related to non-functional software qualities plotted over time, continuous topics are trends that occur across time}
  \label{fig:mysql}
\end{figure*}

\shrink
\subsection{Comparing MaxDB and MySQL}

@@

\emph{Do label frequencies change over time?} Yes, MySQL's label
frequencies decreased as it got older. Usability and reliability
labels became more and more infrequent as it matured. Maintainability topics
became more prevalent as MaxDB matured.

\emph{Do the different projects differ in their relative topic
  interest?} Yes they do, MySQL 3.23 had more proportionally more
functionality labelled topics while MaxDB had proportionally more
efficiency and portability related topics.

@@

\shrink
\subsection{Annotation observations}

@@

\shrink
\subsection{Effectiveness}

@@

\shrink
\subsection{Threats to validity}

@@

\shrink
\section{Related work}

@@

\shrink
\section{Conclusions and future work}

@@

%\section{Acknowledgements and appendix}
Our data and scripts are available at \url{http://softwareprocess.es/nomen/}
